QPLIB: A Library of Quadratic Programming Instances

### 2018-08-08 QPLIB-3.tex
Update QPLIB-3.tex

In this section we describe the procedure we adopted to gather the
instances. In January $2014$, we issued an online call for instances
using main international mailing lists of the mathematical
optimization and numerical analysis communities, reaching in this way
a \textcolor{red}{large set of possibly interested} researchers and practitioners.
The call remained open for ten months, during which we received a large
number of contributions of different nature. The instances we gathered
come both from theoretical studies as well as from real-world
applications.

@@

Other quadratic instances were found in online libraries devoted to
specific QP problems as Max-Cut, Quadratic Assignment, Portfolio
Optimization, and several others. \textcolor{red}{In addition, we mention
that other generic libraries exist, e.g., 
Conic library CBLIB (\url{http://cblib.zib.de}) and 
MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}

@@

\textcolor{red}{We now analyze the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{lllr}
\toprule
Obj. Fun. & Variables & Constraints & \#\\
\cmidrule(lr){1-4}
%
\multirow{5}*{Linear}
          & \multirow{1}*{Binary}
%                    & None      &   \\
%          &         & Linear    &  \\
                    & Quadratic &   9 \\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{2}*{Mixed}
                    & Convex    &   \textcolor{red}{14}\\[1.2 ex]
          &         & Quadratic &  \textcolor{red}{134}\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{Integer}
%                    & Linear    &    \\
                   & Quadratic &    2\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{General}
%                    & Linear    &    \\
                   & Quadratic &    3\\[1.2 ex]
\cmidrule(lr){1-4}
%\multirow{5}*{Linear}
%          & Binary  & Quadratic &   9\\
%          & \multirow{2}*{Mixed}
%                    & Convex    &   2\\
%          &         & Quadratic &  118\\
%%          & \multirow{2}*{Integer}
%%                    & Convex    &  15 \\
%%          &         & Quadratic &   5 \\
%          & {General} & Quadratic &   3 \\
%\hline
\multirow{1}*{Convex (if min)}
          & Binary  & Linear    &  \textcolor{red}{5} \\[1.2 ex]
\cmidrule(lr){2-4}
\multirow{1}*{or}
          & \multirow{2}*{Mixed}
                    & Linear    &   12\\[1.2 ex]
\multirow{1}*{Concave (if max)}
          &         & Quadratic &    \textcolor{red}{6}\\[1.2 ex]
%          & General & Linear    &    \\
%\cmidrule(lr){2-4}
%          & \multirow{1}*{Integer}
%                    & Linear    &   1\\[1.2 ex]
%\hline
\cmidrule(lr){1-4}
\multirow{6}*{Quadratic}
          & \multirow{3}*{Binary}
                    & None      &   23\\[1.2 ex]
          &         & Linear    &  \textcolor{red}{91}\\[1.2 ex]
          &         & Quadratic &   5 \\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{2}*{Mixed}
                    & Linear    &   11\\[1.2 ex]
          &         & Quadratic &    1\\[1.2 ex]
\cmidrule(lr){2-4}
          & Integer & Linear    &    2\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{General}
                    & Quadratic    &    1\\[1.2 ex]
%          &         & Quadratic &    \\
\hline
Total     &         &           & \textcolor{red}{319}\\
%
\bottomrule
\end{tabular}
\caption{Classification of the final set of discrete instances}
\label{tab:DD}
\end{table}

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{llr}
\toprule
Obj. Fun. & Constraints & \#\\
\cmidrule(lr){1-3}
%
\multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
                         & Quadratic &   \textcolor{red}{52}\\[1.2 ex]
\cmidrule(lr){1-3}
\multirow{1}*{Convex (if min)}
                         & Box       &   3 \\[1.2 ex]
\multirow{1}*{or}
                         & Linear    &   \textcolor{red}{16}\\[1.2 ex]
                        % & Convex    &    2\\[1.2 ex]
\multirow{1}*{Concave (if max)}
                         & Quadratic &    \textcolor{red}{11}\\[1.2 ex]
\cmidrule(lr){1-3}
\multirow{3}*{Quadratic}
                         & Linear    &   6\\[1.2 ex]
                         & Convex    &   \textcolor{red}{3}\\[1.2 ex]
                         & Quadratic &   \textcolor{red}{30}\\[1.2 ex]
\hline
Total                    &           & \textcolor{red}{134} \\
%
\bottomrule
\end{tabular}
\caption{Classification of the final set of continuous instances}
\label{tab:CC}
\end{table}

@@

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constrains
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. \textcolor{red}{The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

@@

On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by 
 percentage of ``problematic'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
we mean \textcolor{red}{eigenvalues below $-10^{-12}$} in case of a minimization problem and \textcolor{red}{eigenvalues above $10^{-12}$} in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``problematic'' eigenvalues. Among the others, a vast
majority has around 50\% of ``problematic'' eigenvalues. However, instances
with high or low percentages of ``problematic'' eigenvalues are present, too.

@@

Beyond the instances, the website provides a rich set of metadata for
each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), \textcolor{red}{the contributor of the instance}, basic properties such as the number of variables and constraints of
different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
problem.
%
In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
Lagrangian function, \textcolor{red}{if the instance size allows}.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
\textcolor{red}{Finally, feasible solution points are provided for most instances.}

\begin{figure}
\centering
\includegraphics[scale=0.5]{{QPLIB_2967.jac}.png} \qquad
\includegraphics{{QPLIB_2967.hess}.png}
\caption{Example for the sparsity pattern of the Jacobian of the constraint functions (left) and
  of \textcolor{red}{the upper-right triangle of} the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
  The gradient of the objective function is displayed as the first row
  of the Jacobian matrix. \textcolor{red}{Non-constant entries are shown in red.}}
\label{fig:sparsitypattern}
\end{figure}

The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, \textcolor{red}{problematic} eigenvalues in~$Q^0$, and density of~$Q^0$.
%
Finally, a statistics page displays diagrams on the composition of the library according to different criteria: the
number of instances according to problem type, variable and constraint types, convexity, problem size, and density.
%
A file containing the relevant metadata for each instance can be downloaded in comma-separated-values (\texttt{csv}) format,
so that researchers can easily compile their own subset of instances according to these statistics.

>>>>

In this section we describe the procedure we adopted to gather the
instances. In January $2014$, we issued an online call for instances
using main international mailing lists of the mathematical
optimization and numerical analysis communities, reaching in this way
a {large set of possibly interested} researchers and practitioners.
The call remained open for ten months, during which we received a large
number of contributions of different nature. The instances we gathered
come both from theoretical studies as well as from real-world
applications.

@@

Other quadratic instances were found in online libraries devoted to
specific QP problems as Max-Cut, Quadratic Assignment, Portfolio
Optimization, and several others. {In addition, we mention
that other generic libraries exist, e.g., 
Conic library CBLIB (\url{http://cblib.zib.de}) and 
MIPLIB~2010 (\url{http://miplib.zib.de/}), to mention just a few.}

@@

{We now analyze the features of the instances selected to be part of the
 library. In Table \ref{tab:DDDDD}, we provide a global overview. The instances have been divided in \textit{continuous} vs \textit{discrete} and \textit{convex} vs \textit{nonconvex}, forming in this way, a  classification of 4 macro categories. As previously mentioned, an instance is classified \textit{discrete} if it contains at least a binary or integer variable, and \textit{continuous} otherwise. On the other hand, an instance is classified as \textit{nonconvex} if the objective function is nonconvex (if minimization) or nonconcave (if maximization) and/or at least one of the constraints is nonconvex, and \textit{convex} otherwise. }

@@

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{lllr}
\toprule
Obj. Fun. & Variables & Constraints & \#\\
\cmidrule(lr){1-4}
%
\multirow{5}*{Linear}
          & \multirow{1}*{Binary}
%                    & None      &   \\
%          &         & Linear    &  \\
                    & Quadratic &   9 \\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{2}*{Mixed}
                    & Convex    &   {14}\\[1.2 ex]
          &         & Quadratic &  {134}\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{Integer}
%                    & Linear    &    \\
                   & Quadratic &    2\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{General}
%                    & Linear    &    \\
                   & Quadratic &    3\\[1.2 ex]
\cmidrule(lr){1-4}
%\multirow{5}*{Linear}
%          & Binary  & Quadratic &   9\\
%          & \multirow{2}*{Mixed}
%                    & Convex    &   2\\
%          &         & Quadratic &  118\\
%%          & \multirow{2}*{Integer}
%%                    & Convex    &  15 \\
%%          &         & Quadratic &   5 \\
%          & {General} & Quadratic &   3 \\
%\hline
\multirow{1}*{Convex (if min)}
          & Binary  & Linear    &  {5} \\[1.2 ex]
\cmidrule(lr){2-4}
\multirow{1}*{or}
          & \multirow{2}*{Mixed}
                    & Linear    &   12\\[1.2 ex]
\multirow{1}*{Concave (if max)}
          &         & Quadratic &    {6}\\[1.2 ex]
%          & General & Linear    &    \\
%\cmidrule(lr){2-4}
%          & \multirow{1}*{Integer}
%                    & Linear    &   1\\[1.2 ex]
%\hline
\cmidrule(lr){1-4}
\multirow{6}*{Quadratic}
          & \multirow{3}*{Binary}
                    & None      &   23\\[1.2 ex]
          &         & Linear    &  {91}\\[1.2 ex]
          &         & Quadratic &   5 \\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{2}*{Mixed}
                    & Linear    &   11\\[1.2 ex]
          &         & Quadratic &    1\\[1.2 ex]
\cmidrule(lr){2-4}
          & Integer & Linear    &    2\\[1.2 ex]
\cmidrule(lr){2-4}
          & \multirow{1}*{General}
                    & Quadratic    &    1\\[1.2 ex]
%          &         & Quadratic &    \\
\hline
Total     &         &           & {319}\\
%
\bottomrule
\end{tabular}
\caption{Classification of the final set of discrete instances}
\label{tab:DD}
\end{table}

\begin{table}
 \centering
 %\scriptsize
 \setlength{\tabcolsep}{18pt}
 \renewcommand \arraystretch{1.1}
\begin{tabular}{llr}
\toprule
Obj. Fun. & Constraints & \#\\
\cmidrule(lr){1-3}
%
\multirow{2}*{Linear}    & Convex    &   13\\[1.2 ex]
                         & Quadratic &   {52}\\[1.2 ex]
\cmidrule(lr){1-3}
\multirow{1}*{Convex (if min)}
                         & Box       &   3 \\[1.2 ex]
\multirow{1}*{or}
                         & Linear    &   {16}\\[1.2 ex]
                        % & Convex    &    2\\[1.2 ex]
\multirow{1}*{Concave (if max)}
                         & Quadratic &    {11}\\[1.2 ex]
\cmidrule(lr){1-3}
\multirow{3}*{Quadratic}
                         & Linear    &   6\\[1.2 ex]
                         & Convex    &   {3}\\[1.2 ex]
                         & Quadratic &   {30}\\[1.2 ex]
\hline
Total                    &           & {134} \\
%
\bottomrule
\end{tabular}
\caption{Classification of the final set of continuous instances}
\label{tab:CC}
\end{table}

@@

We now report some graphs that help in illustrating the main features
of the instances. In Figure~\ref{fig:distribution}~(left) we plot the number
of variables (horizontal axis) versus the number of constrains
(vertical axis), both in logarithmic scale. Continuous instances are
marked with ``$+$'', and discrete ones with ``$\times$''. The figure shows
that the library contains a quite diverse set of instances in
terms of number of variables and constraints. {The record on the maximal number of variables and constraints (both $\approx 1,000,000$) is set by the instances QPLIB\_8547 and QPLIB\_9008.}
Figure~\ref{fig:distribution}~(right) plots the number of nonzero elements in the gradient of the objective function and the Jacobian and the number of these nonzeros corresponding to nonlinear variables, that is, it counts the appearances of variables in objectives and constraints and how often such an appearance is in a quadratic term.

@@

On the theme of nonconvexity, Figure~\ref{fig:pic_neg_eig} (left) focuses on
the instances with a quadratic objective function, ordered by 
 percentage of ``problematic'' eigenvalues in the Hessian $Q^0$ (vertical axis), by which
we mean {eigenvalues below $-10^{-12}$} in case of a minimization problem and {eigenvalues above $10^{-12}$} in case of a maximization problem.
The instances are mostly clustered around two values. About 25\% of the instances have a convex (if minimization) or concave (if maximization) objective function,
i.e., they have 0\% of ``problematic'' eigenvalues. Among the others, a vast
majority has around 50\% of ``problematic'' eigenvalues. However, instances
with high or low percentages of ``problematic'' eigenvalues are present, too.

@@

Beyond the instances, the website provides a rich set of metadata for
each instance: the three letter problem classification (as described in \S\ref{subsec:final set}), {the contributor of the instance}, basic properties such as the number of variables and constraints of
different types, the sense and convexity/concavity of the objective function, and information on the nonzero structure of the
problem.
%
In addition, we display a visualization of the sparsity patterns of the Jacobian and the Hessian matrix of the
Lagrangian function, {if the instance size allows}.  In the plots of the Jacobian nonzero pattern, the linear and nonlinear entries are distinguished
by color.  Figure \ref{fig:sparsitypattern} shows an example for instance QPLIB\_2967.
{Finally, feasible solution points are provided for most instances.}

\begin{figure}
\centering
\includegraphics[scale=0.5]{{QPLIB_2967.jac}.png} \qquad
\includegraphics{{QPLIB_2967.hess}.png}
\caption{Example for the sparsity pattern of the Jacobian of the constraint functions (left) and
  of {the upper-right triangle of} the Hessian of the Lagrangian function (right) for instance QPLIB\_2967.
  The gradient of the objective function is displayed as the first row
  of the Jacobian matrix. {Non-constant entries are shown in red.}}
\label{fig:sparsitypattern}
\end{figure}

The entire set of instances can be explored in a searchable and sortable table of selected instance features: problem
classification, convexity of the continuous relaxation, number of (all, binary, integer) variables, (all, quadratic) constraints, nonzeros, {problematic} eigenvalues in~$Q^0$, and density of~$Q^0$.
%
Finally, a statistics page displays diagrams on the composition of the library according to different criteria: the
number of instances according to problem type, variable and constraint types, convexity, problem size, and density.
%
A file containing the relevant metadata for each instance can be downloaded in comma-separated-values (\texttt{csv}) format,
so that researchers can easily compile their own subset of instances according to these statistics.
